---
title: "Class 09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
read.csv("Data Export Summary.csv")
  

```
```{r}
163351/187423 * 100
10139/187423 * 100
```
>Q1: 87.15633%
>Q2: 5.409688%
>Q3. 3819
>Q4. The 2 hydrogen atoms show up as charges.

```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
```
>Q10. msa
>Q11. devtools
>Q12. TRUE

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```

>Q13. 214


```{r}
b <- blast.pdb(aa)
hits <- plot(b)
```


```{r}
head(hits$pdb.id)
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
library(muscle)
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```












## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
