---
title: "class 14"
author: "Ramola Baviskar (PID A12228297)"
date: "3/7/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>Important vaccination data

```{r}
vax <- read.csv("vaccinedata.csv")
head(vax)
```
>Q1. persons_fully_vaccinated
>Q2. zip_code_tabulation_area
>Q3. 2021-01-05
>Q4. 2022-02-22


```{r}
skimr::skim(vax)
```

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

```{r}
nrow(vax)
sum(is.na(vax$persons_fully_vaccinated)) / nrow(vax)
```

>Q5. 9
>Q6. 18338
>Q7. ~17%
>Q8. Some counties may not have collected this data yet or haven't yet submitted it.

```{r}
library(lubridate)
vax$as_of_date <- ymd(vax$as_of_date)
today <- today() - vax$as_of_date
now <- vax$as_of_date[nrow(vax)] - vax$as_of_date
```


```{r}
date <- vax[, c("as_of_date")]
length(unique(date))
```

>Q9. 5=425-420
>Q10. 61

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109"))
```

```{r}
sd <- vax[ "San Diego" , ]
library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
sd.10 <- filter(vax, county == "San Diego" & age5_plus_population > 10000)
```

```{r}
sd <- filter(vax, county == "San Diego")
code <- sd[, c("zip_code_tabulation_area")]
length(unique(code))
```

>Q11. 107

```{r}
sd12 <- filter(vax, county == "San Diego" & age12_plus_population > 60000)
sd12[, c("age12_plus_population")]
which.max("age12_plus_population")
```

>Q12. 92154

```{r}
fullyvaxxed <- filter(vax, county == "San Diego", as_of_date == "2022-02-22")
nrow(fullyvaxxed)
percentvaxxed <- fullyvaxxed[, c("percent_of_population_fully_vaccinated")]
percentvaxxedomit <- na.omit(percentvaxxed)
sum(percentvaxxedomit)/nrow(fullyvaxxed) * 100
```

>Q13. 69.75742%

```{r}
library(ggplot2)
```

>Q14.

```{r}
ggplot(data=fullyvaxxed, aes(percent_of_population_partially_vaccinated)) + geom_histogram()
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>Q15. 
>Q16.

```{r}
lineplot <- ggplot(ucsd) +
  aes(x=as_of_date, y=percent_of_population_fully_vaccinated) +
  geom_line() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") 
lineplot + geom_hline(yintercept=0.733385, linetype="dashed", 
                color = "red", size=1)
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 & as_of_date == "2022-02-22")
mean(vax.36[, c("percent_of_population_fully_vaccinated")])
min(vax.36[, c("percent_of_population_fully_vaccinated")])
median(vax.36[, c("percent_of_population_fully_vaccinated")])
max(vax.36[, c("percent_of_population_fully_vaccinated")])
quantile(vax.36[, c("percent_of_population_fully_vaccinated")])
```

>Q17. Min:0.3881090 1st Q:0.6539015 Median:0.7332750 3rd Q:0.8027110 Max:1.0000000 Mean:0.733385
>Q18. 

```{r}
ggplot(data=ucsd, aes(percent_of_population_partially_vaccinated)) + geom_histogram()
```

```{r}
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

>Q19: 92040: 0.551304
      92019: 0.723044 Both below average.
      
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
```

>Q20.

```{r}
ggplot(vax.36.all) +
  aes(x=as_of_date, y=percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rates Across CA",
       subtitle="Only areas with a population above 36,000 are shown") +
  geom_hline(yintercept = 0.733385, linetype="dashed")
```

>Q21. I have mixed feelings about it, since there are parts of the state with quite low vaccination rates as of just a few days ago; anyone who travels there has a greater risk of coming into contact with the virus. 





