---
title: "Class6:Rfunctions"
author: "RamolaBaviskar (PID A12228297)"
date: "2/3/2022"
output:
  pdf_document: default
  html_document: default
---

Today we will explore R functions.

We will start with calculating a grade for these example students.
```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

We can use the mean() function to get the average.
```{r}
mean(student1)

```

```{r}
mean(student2, na.rm=TRUE)
```

How does the is.na() function work? Let's try it on student 2. 
```{r}
student2
is.na(student2)

```

We can use this result to get at our NA values, the TRUE positions.
```{r}
student2[is.na(student2)] <- 0
student2
mean(student2)
```
```{r}
student3[is.na(student3)] <- 0
student3
mean(student3)
```

Now we'll find the lowest score.
```{r}
min(student1)
```


```{r}
student1
min(student1)
which.min(student1)
x <- student1[-which.min(student1)]
mean(x)
```

Now we can turn this into our first function, called 'grade()'.
All R functions have 3 things: a name (grade), input arguments (scores), and body (does the work).

```{r}
#first set NA values to zero

x[is.na(x)] <- 0
#remove lowest score and calculate mean
x <- x[-which.min(x)]
mean(x)
```
```{r}
#first set NA values to zero
grade <- function(x) {
  x[is.na(x)] <- 0
  #remove lowest score and calculate mean
  mean(x[-which.min(x)])
}
```

Let's try it
```{r}
grade(student3)

```

Q1. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]


```{r}
gradebook <- read.csv("https://tinyurl.com/gradeinput", row.names=1)
head(gradebook)
```
 
 We can use the 'apply()' function to grade the whole class.
 
```{r}
apply(gradebook, 1, grade)
scores <- apply(gradebook, 1, grade)
```
 
Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]

```{r}
which.max(scores)
```

Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]
```{r}
  hw.mean <- (apply(gradebook, 2, mean, na.rm = TRUE))
  which.min(hw.mean)
```

```{r}
boxplot(gradebook)
```
 
  According to the boxplot, HW2 has the lowest average but HW3 has the smallest spread and the lowest overall scores. Therefore, HW3 was the toughest on students.





Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

  





